### 2021-06-13 letter/responseLetter.tex

\begin{itemize}

\item It is not clear how this paper can be considered an extension of the DroidXP paper presented in SCAM 2020.

\vspace{0.2cm}

{\color{blue}{\bf Answer:} In the introduction, we rephrase paragraphs 5 and 6 to better clarify how our paper is an extension of DroidXP paper from SCAM 2020.}

@@

\vspace{0.2cm}

\item ``The authors state that dynamic analysis exceeds static analysis on mining Android sandboxes because the
analysis must often assume that additional behaviors are possible than actually would be.''
Please explain

\vspace{0.2cm}

{\color{blue}{\bf Answer:} We explained better the sentence at Introduction, paragraph 3, where we add the follow text: ``That is because many times codes can be interpreted, decrypted, or even downloaded at runtime. In these cases, the static analysis could be ineffective for mining sandboxes, since that if static analysis determines that some behavior will not occur, it no longer will be checked at runtime.''}

@@

\vspace{0.2cm}

\item ``We presented and evaluated DroidXP in a conference paper [11]''
What is DroidXP? It should be introduced to a reader who is not familiar.

\vspace{0.2cm}

{\color{blue}{\bf Answer:} We agree with that. To mitigate this issue, we introduced better DroidXP, contextualizing why its development was important for our study. Text improvement is present in the paragraph 5 of introduction}

@@

\vspace{0.2cm}

\item Google play - Google Play
* java - Java / android - Android
* ``Another approach to rooting''
What is rooting?

\vspace{0.2cm}

{\color{blue}{\bf Answer:} Sorry about these typos. We corrected all of them. Thank one more time. With respect to ``rooting'': It's a process performed on Android devices like smartphones or tablets, which allows users of the Android mobile operating system, to execute privileged commands that are typically unavailable at its default configuration. The term comes from the fact that Android system is based on Linux, and this system has a (superuser), called root, which has access to administrative permissions as on Linux. We insert a footnote with a link that better explain this term at page 4.\newline
(https://en.wikipedia.org/wiki/Rooting(Android))
}

@@

\vspace{0.2cm}

\item Briefly explain the taint analysis
``Taint analysis is a special type of static analysis that enables tracking of sensitive data within programs''

\vspace{0.2cm}

{\color{blue}{\bf Answer:} We agree that taint analysis explanation was very briefly. We reviewed our Background and related works Section, to better explain taint analysis from paragraph 8. We explained better the concept of ``source'' and ``sink'', and present how the analysis identifies sensitive information leakage, detecting taint flow between “source” and “sink”. To improve understanding, we also present a simple Listing (Listing 1), that present a simple data leakage example through data flows within source code Java. }

@@

{\color{blue}{\bf Answer:} We agree that the sentence was messy. We rephrased the sentence to clarify the first paragraph of Section 4.1. ``Then we repeated the process, however this time we isolate the effect of the static analysis component of DroidFax.''}

\vspace{0.2cm}

\item Section 4.1 can be written more concisely.
* It is not clear what is the value of the malware discussion between pages 13 and 15. It may be removed.

what  are  thecharacteristics  of  the  malwares  that  have  (not)  been  identified  using  the  mining  sandbox  approach

Our goal here is to provide a lower-level intuition about the classes of malware the mining sandbox approach is able to detect.

@@

\vspace{0.2cm}

{\color{blue}{\bf Answer:} Thanks for raising this issue. Actually our purpose was to present the characteristics of malware that have or have not been identified using the mining sandox approach. Our goal is to provide a lower-level intuition about the classes of malware, the mining sandbox approach is good enough, and which ones it fails. We rephrased the section 4.1 at this point, to clarify this issue about malware discussion.}

@@

\vspace{0.2cm}

\item There are several other typos and grammatical errors.

\vspace{0.2cm}

{\color{blue}{\bf Answer:} Sorry about these typos and grammar. We revised all the text, and corrected this problem like at page 9 (does not computed – does not compute), and page 10 (of our experimental – of our experiments), and some other mistakes that went unnoticed.}

### 2021-06-13 letter/responseLetter.tex

\vspace{0.2cm}

\item In the second place, I would recommend to include more details on the experiments conducted. This is especially 
true for the second study, where Section 3.3 is a bit shallow. The three steps of the methodology are indeed shortly 
described without providing enough details, hence possibly precluding verifiability and replicability. In addition, the 
methodology is not linked to the research questions: I would recommend to restructure Section 3 in order to have 
data extraction processes and data analysis procedures for each research question.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We changed Section 3.3 to detail our approach better. We also created a picture that can help understand the methodology of our second study.}

@@

\begin{itemize}

\item It is not clear how this paper can be considered an extension of the DroidXP paper presented in SCAM 2020.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} In the introduction, we rephrase paragraphs 5 and 6 to better clarify how our paper is an extension of DroidXP paper from SCAM 2020.}

@@

\vspace{0.2cm}

\item ``The authors state that dynamic analysis exceeds static analysis on mining Android sandboxes because the
analysis must often assume that additional behaviors are possible than actually would be.''
Please explain

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We explained better the sentence at Introduction, paragraph 3, where we add the follow text: ``That is because many times codes can be interpreted, decrypted, or even downloaded at runtime. In these cases, the static analysis could be ineffective for mining sandboxes, since that if static analysis determines that some behavior will not occur, it no longer will be checked at runtime.''}

@@

\vspace{0.2cm}

\item ``We presented and evaluated DroidXP in a conference paper [11]''
What is DroidXP? It should be introduced to a reader who is not familiar.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We agree with that. To mitigate this issue, we introduced better DroidXP, contextualizing why its development was important for our study. Text improvement is present in the paragraph 5 of introduction}

@@

\vspace{0.2cm}

\item Google play - Google Play
* java - Java / android - Android
* ``Another approach to rooting''
What is rooting?

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Sorry about these typos. We corrected all of them. Thank one more time. With respect to ``rooting'': It's a process performed on Android devices like smartphones or tablets, which allows users of the Android mobile operating system, to execute privileged commands that are typically unavailable at its default configuration. The term comes from the fact that Android system is based on Linux, and this system has a (superuser), called root, which has access to administrative permissions as on Linux. We insert a footnote with a link that better explain this term at page 4.\newline
(https://en.wikipedia.org/wiki/Rooting(Android))
}

@@

\vspace{0.2cm}

\item Briefly explain the taint analysis
``Taint analysis is a special type of static analysis that enables tracking of sensitive data within programs''

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We agree that taint analysis explanation was very briefly. We reviewed our Background and related works Section, to better explain taint analysis from paragraph 8. We explained better the concept of ``source'' and ``sink'', and present how the analysis identifies sensitive information leakage, detecting taint flow between “source” and “sink”. To improve understanding, we also present a simple Listing (Listing 1), that present a simple data leakage example through data flows within source code Java. }

@@

\vspace{0.2cm}

\item Section 4.1 can be written more concisely.
* It is not clear what is the value of the malware discussion between pages 13 and 15. It may be removed.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Thanks for raising this issue. Actually our purpose was to present the characteristics of malware that have or have not been identified using the mining sandox approach. Our goal is to provide a lower-level intuition about the classes of malware, the mining sandbox approach is good enough, and which ones it fails. We rephrased the section 4.1 at this point, to clarify this issue about malware discussion.}

@@

\vspace{0.2cm}

\item There are several other typos and grammatical errors.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Sorry about these typos and grammar. We revised all the text, and corrected this problem like at page 9 (does not computed – does not compute), and page 10 (of our experimental – of our experiments), and some other mistakes that went unnoticed.}

### 2021-06-14 letter/responseLetter.tex

\documentclass[12pt,english]{scrlttr2}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[]{babel}

\newkomavar{fromplace}
\setkomavar{fromname}{Francisco Handrick da Costa}
\setkomavar{fromaddress}{Fairytail-Lane 1}
\setkomavar{fromplace}{12345 Cologne}

\firsthead{%
\begin{tabular}{p{0.7\textwidth}p{0.3\textwidth}}
& \usekomavar{fromname} \newline \usekomavar{fromaddress}  \newline \usekomavar{fromplace} 
\end{tabular}
}

% \documentclass{letter}

% \usepackage{natbib}
\usepackage{hyperref}
\usepackage{xcolor}
\DeclareOldFontCommand{\bf}{\normalfont\bfseries}{\mathbf}

% \signature{Francisco Handrick Tomaz da Costa (on behalf of the authors of the paper)}

%\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil \\\\ Software Technology Group
% \\ Technical University of Darmstadt \\ Germany}

@@

\vspace{0.2cm}

\item The proposed DroidXP benchmark seems not to be available, which is a pity since it might be very useful for other 
researchers. I believe the paper may have a stronger impact if the tool would be available. The same is true for the 
data and scripts used for the experiments. I would appreciate if the paper may have an online repository making all 
data and tools available.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Thanks for this comment. We make avaiable our tool, datasets and
  scripts in the following links:

  \begin{itemize}
    \item \href{https://github.com/droidxp/benchmark}{GitHub repository}.
    \item \href{https://htmlpreview.github.io/?https://github.com/droidxp/paper-replication-package/blob/master/replication.html}{datasets and scripts}
  \end{itemize}

\vspace{0.2cm}\newline
At Section 1 (Introduction) we highlight the existence of these 2 repositories that can be useful for other researchers interested in our study.}

@@

\vspace{0.2cm}

\item The significance of this work is questionable. The first study is a mere replication of the existing work [2], except, disabling 
part of the framework. The finding is that DroidFaX is important part for DroidXP. Well, your proved literature's correctness. 
It would be more impressive the other way round.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We respectfully disagree with this opinion. The relevance of replication in software enginer has gained much more attention in recent years (including recommendations for that~\cite{role-of-replication,standards}),
  not only to validate past results~\cite{msr-replication} but also to generate new insights
  from previous research results~\cite{} or even refute past conclusions~\cite{topas-replication}.
  Indeed, we carried out a non-exact replication of~\cite{}, and the results of our study
  sheds light that the performance of the dynamic analysis tools explored in~\cite{} is
  actually different from the previous report. We also agree with the perspective of Shullet al.~\cite{},
  who argue that an exact replication that achieves the same conclusions of an original work
  is still valuable to generalize the results of a research. Nonetheless, here we go beyond that,
  because we give evidence that the performance of the dynamic analysis tools in the previous research
  work had been overestimated. To make this point clear, we changed \ldots}

@@

\vspace{0.2cm}

\item I find what's written in Finding 2 is not exciting. It is intuitive based on the results of [1] to know that a comprehensive 
comparison combining multiple dynamic tools would improve the performance.

\vspace{0.2cm}

{\color{blue}{\bf Answer.}The \emph{Finding 2 box} summarizes that we achieve the best performance on mining sandboxes when combining Humanoid, Monkey and DroidBot. Adding DroidMate to this configuration does not improve the overall performance. This result is novel and different from the previous research. We have made a slight change in this particular box to stress this point.}

@@

\vspace{0.2cm}

\item The title itself is inaccurate. You removed DroidFax from the replicated work[1]. But it is not the only static analysis part. 
Actually, many dynamic testing tools use static analysis to better guide their testing strategy.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We agree. Nonetheless, DroidFax is not part of the dynamic analysis tools. According to Bao et al., their goal on using DroidFax was to instrument the APKs with the intent of collecting calls to sensitive APIs. However, the use of DroidFax in their study actually lead to an increase in the peformance of dynamic analysis tools for mining sandboxes. We did not change any internal component of the dynamic analysis tools, which indeed might include static analysis components. We review the title to mitigate this problem.}

@@

\vspace{0.2cm}

\item Typos:
1. Page 3: to explains -- to explain.
2. Page 3: BLL-Study study -- BLL-Study.
3. Page 8, does not computed -- does not compute.
4. Page 9, of our experimental -- of our experiments.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Sorry about these typos and grammar issues. We corrected all of them. Thanks.}

### 2021-06-14 background.tex

Nowadays, malware becomes more stealthy and hackers learn how to avoid anti-virus signature checks, by obfuscating the native code~\cite{DBLP:journals/corr/abs-2002-04540}
%\kn{Here we should cite the work of String hound (https://dl.acm.org/doi/10.1145/3320269.3384745)}
, calling Java libraries, or even by creating root exploits, since the native code is allowed to directly make syscalls. Another approach to rooting is making a side attack aimed at a benign app to make a syscall, exploiting its native code to get root access.

>>>>

Nowadays, malware becomes more stealthy and hackers learn how to avoid anti-virus signature checks, by obfuscating the native code~\cite{DBLP:journals/corr/abs-2002-04540}
%\kn{Here we should cite the work of String hound (https://dl.acm.org/doi/10.1145/3320269.3384745)}
, calling Java libraries, or even by creating root exploits, since the native code is allowed to directly make syscalls. Another approach to rooting\footnote{https://en.wikipedia.org/wiki/Rooting(Android)} is making a side attack aimed at a benign app to make a syscall, exploiting its native code to get root access.

### 2021-06-14 introduction.tex

Jamrozik et al.~\cite{DBLP:conf/icse/JamrozikSZ16} argue in favor of dynamic analysis for mining sandboxes, instead of using static analysis.
The authors state that dynamic analysis exceeds static analysis on mining Android sandboxes because the analysis must often
assume that additional behaviors are possible than actually would be.
%\kn{Ok. In the first paragraph we discuss static analysis in general to detect vulnerabilities and now without mentioning it prior we seem to mention static analysis can also mine sandboxes. The overall flow of definitions needs some thought}
Otherwise, static analysis might determine that some behavior is impossible and therefore no longer needs to be checked at runtime. In addition,
code that uses dynamic features (such as reflection) poses additional challenges to static analysis algorithms---even though \emph{dynamic features}
of programming languages are often used to introduce malicious behavior. Even though these claims
seem reasonable, previous research do not present empirical assessments about the limitations of static analysis to mine sandboxes.
Consequently, it is not clear whether and how both approaches (dynamic and static analysis) could complement each other in the process of mining Android sandboxes.

@@

Our goal in this paper is to understand how static analysis
algorithms might complement dynamic analysis approaches in the process of mining Android sandboxes. Here we
present the results of two empirical studies that aim to
address our research goal. In the
first study, we conduct a replication of the \blls, though our research differs from the original work in three aspects: (a)
we isolate the effect of the static analysis component of DroidFax, (b) we included a recent dynamic analysis tool for testing Android apps,
and (c) we extend the execution time and the number of executions of each test case generator tool for all apps
in our dataset. Differently, in the second study we explore how a more advanced static analysis approach
(based on taint analysis) can complement dynamic analysis in the task of mining sandboxes---using
the FlowDroid~\cite{DBLP:conf/pldi/ArztRFBBKTOM14} infrastructure for analysing Android apps.

We presented and evaluated DroidXP in a conference paper~\cite{DBLP:conf/scam/CostaMCMVBC20},
and here we discuss the results of a broader investigation
that also uses DroidXP to explain the performance of static and dynamic analysis
approaches for mining sandboxes. Altogether, this paper brings the following contributions:

@@

\item A broad comprehension about the role of static analysis tools for mining
  sandboxes, showing that we can benefit from using both static and dynamic
  analysis for detecting Android malwares.

@@

\item A replication package of our study that is available on-line, including
  scripts for statistic analysis \footnote{https://htmlpreview.github.io/?https://github.com/droidxp/paper-replication-package/blob/master/replication.html}
  and a benchmark for reproducing and extending our study.\footnote{https://github.com/droidxp/benchmark} 
\end{itemize}

>>>>

Jamrozik et al.~\cite{DBLP:conf/icse/JamrozikSZ16} argue in favor of dynamic analysis for mining sandboxes, instead of using static analysis.
The authors state that dynamic analysis exceeds static analysis on mining Android sandboxes because the analysis must often
assume that additional behaviors are possible than actually would be. That is because many times codes can be interpreted, decrypted, or even downloaded at runtime. In these cases, the static analysis could be ineffective for mining sandboxes, since that if static analysis determines that some behavior will not occur, it no longer will be checked at runtime. In addition, code that uses dynamic features (such as reflection) poses additional challenges to static analysis algorithms---even though \emph{dynamic features} of programming languages are often used to introduce malicious behavior. Even though these claims seem reasonable, previous research do not present empirical assessments about the limitations of static analysis to mine sandboxes. Consequently, it is not clear whether and how both approaches (dynamic and static analysis) could complement each other in the process of mining Android sandboxes.

@@

Here we replicated \blls, benefiting from our benchmark tool called DroidXP. We developed DroidXP to help researchers and practitioners (including ourselves) to integrate test case generation tools and compare their performance on mining sandboxes at Android platform. The repository for DroidXP is available at GitHub\footnote{https://github.com/droidxp/benchmark} where offers direct access to all source code, installation process, dependencies, and how to run. We presented and evaluated DroidXP in a conference paper~\cite{DBLP:conf/scam/CostaMCMVBC20}, and here we discuss the results of a broader investigation that also uses DroidXP to explain the performance of static and dynamic analysis approaches for mining sandboxes. Our goal is to understand how static analysis algorithms might complement dynamic analysis approaches in the process of mining Android sandboxes. This work is an extension of SCAM 2020 paper, since that for our actual study, we had to adapt DroiXP to allow it to isolate the effects of static analysis component of DroidFax, and take into account this effect at our analysis.

Our research work presents the results of two empirical studies that aim to
address our research goal. In the
first study, we conduct a replication of the \blls, as presented at SCAN 2020 conference~\cite{DBLP:conf/scam/CostaMCMVBC20}, though our research differs from the original work in three aspects:

\begin{enumerate}[(a)]
  \item We isolate the effect of the static analysis component of DroidFax.
   \item We included a fake test
case generation tool (named \joke). \joke simulates a test tool that does not execute Android apps during DroidXP execution, disregarding the effects of dynamic analysis, and computing the results with only static analysis DroidFax component.
\item We extend the execution time and the number of executions of each test case generator tool for all apps
in our dataset.

\end{enumerate}

Differently, in the second study we explore how a more advanced static analysis approach
(based on taint analysis) can complement dynamic analysis in the task of mining sandboxes---using
the FlowDroid~\cite{DBLP:conf/pldi/ArztRFBBKTOM14} infrastructure for analysing Android apps.

Altogether, this paper brings the following contributions:

@@

\item A broad comprehension about the role of static analysis tools for mining
  sandboxes, showing that we can benefit from using both static and dynamic
  analysis for detecting Android malware.

@@

\item A replication package of our study that is available online, including
  scripts for statistic analysis \footnote{https://htmlpreview.github.io/?https://github.com/droidxp/paper-replication-package/blob/master/replication.html}
  and a benchmark for reproducing and extending our study.
\end{itemize}
